---
title: "First post"
author: "Kevin Chang"
date: 2021-01-05T12:09:13-06:00
categories: R
tags:
- R Markdown
- plot
- regression
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(httr)
library(lubridate)
library(tidyverse)

```

This blog post is to demonstrate how to extract the latest data set from a github repository and generate bar chart using ggplot2 R package. 

# Link to the github repository

```{r link_git}
req <- GET("https://api.github.com/repos/CSSEGISandData/COVID-19/git/trees/master?recursive=1")

filelist <- unlist(lapply(content(req)$tree, "[", "path"), use.names = F)

filelist <- 
  grep("csse_covid_19_data/csse_covid_19_daily_reports/([0-9])",
     filelist, value = TRUE)

```

All file names are stored in the `filelist` object, which looks something like below

```{r, echo = FALSE}
head(filelist, 3)
```

# Extract the latest date

We can get the list of dates from the file names above by 

```{r, eval=FALSE}
basename(filelist)
```

```{r, echo=FALSE}
head(basename(filelist), 3)
```

Since these are still as character, we can convert these file names to `dates` by using `mdy()` function in the `lubridate` R package, which parse dates with, **m**onth, **d**ay and **y**ear components (in this order), i.e. 

```{r, eval=FALSE}
mdy(basename(filelist))
```

```{r, echo=FALSE}
head(mdy(basename(filelist)), 3)
```

Thus, we can get the lastest date from the avalible data set in the reposiity using the `max()` function across all the dates, i.e.

```{r}
dates <-  mdy(basename(filelist))

max(dates, na.rm = TRUE)
```

We can get the previous date from this laest date by taking away one, i.e. 

```{r}
max(dates, na.rm = TRUE) - 1
```



# Extract the latest data set 

Once we obtain the latest dates, we can use this as the name of the file names to read-in the latest COVID-19 data. 

```{r}
latest_daily_cases <-
  read_csv(
    paste0(
      "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/",
      filelist[which(dates == max(dates, na.rm = TRUE))]
    )
  )


```

```{r}
previous_daily_cases <-
  read_csv(
    paste0(
      "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/",
      filelist[which(dates == (max(dates, na.rm = TRUE) - 1))]
    )
  )
```

# Tidy up the date

Current data set has broken down into different states, since we are only interest the records of different countries, we will need to aggregate these number accordingly, i.e. 

```{r}
latest_daily_cases_total <-
  latest_daily_cases %>%
  select(Country_Region, Confirmed:Active) %>%
  rename(Country = Country_Region) %>%
  pivot_longer(Confirmed:Active, names_to = "Type", values_to = "Count") %>%
  group_by(Country, Type) %>%
  summarise(Count_lastest = sum(Count, ne.rm = TRUE), .groups = "drop")

previous_daily_cases_total <-
  previous_daily_cases %>%
  select(Country_Region, Confirmed:Active) %>%
  rename(Country = Country_Region) %>%
  pivot_longer(Confirmed:Active, names_to = "Type", values_to = "Count") %>%
  group_by(Country, Type) %>%
  summarise(Count_previous = sum(Count, ne.rm = TRUE), .groups = "drop")
```


Combine these two data set and compute the differences for plotting

```{r}
daily_cases_total <- 
  latest_daily_cases_total %>% 
  left_join(previous_daily_cases_total) %>% 
  mutate(Count_change = Count_lastest - Count_previous)
```

# Table outputs


## Overall 

```{r}
daily_cases_total %>%
  group_by(Type) %>%
  summarise("Lastest" = sum(Count_lastest, na.rm = TRUE), 
            "Changes" = sum(Count_change, na.rm = TRUE), 
            .groups = "drop") %>% 
  arrange(-Lastest) %>%
  gt::gt() %>% 
  gt::fmt_number(2:3, suffixing = TRUE)
```

## US 

```{r}
daily_cases_total %>% 
  filter(Country == "US") %>%
  group_by(Type) %>%
  summarise("Lastest" = sum(Count_lastest, na.rm = TRUE), 
            "Changes" = sum(Count_change, na.rm = TRUE), 
            .groups = "drop") %>% 
  arrange(-Lastest) %>%
  gt::gt() %>% 
  gt::fmt_number(2:3, suffixing = TRUE)
```

## United Kingdom 

```{r}


daily_cases_total %>% 
  filter(Country == "United Kingdom") %>%
  group_by(Type) %>%
  summarise("Lastest" = sum(Count_lastest, na.rm = TRUE), 
            "Changes" = sum(Count_change, na.rm = TRUE), 
            .groups = "drop") %>% 
  arrange(-Lastest) %>%
  gt::gt() %>% 
  gt::fmt_number(2:3, suffixing = TRUE)
```


# Bar chart

## Confirmed Cases

```{r}
top <- 20

max_y <-
  daily_cases_total %>%
  filter(Type == "Confirmed") %>% pull(Count_lastest) %>% max(na.rm = TRUE)

daily_cases_total %>% 
  filter(Type == "Confirmed") %>% 
  arrange(-Count_lastest) %>% 
  mutate(Country = forcats::fct_reorder(Country, Count_lastest))%>% 
  slice_head(n = top) %>% 
  ggplot(aes(x = Country, y = Count_lastest, 
             label = paste0(round(Count_lastest/1e6,1), "M", 
                            "(+", round(Count_change/1e3, 1), "K)"))) +
  geom_col() +
  geom_text(hjust = 0, fontface = "bold") + 
  scale_y_continuous(labels = scales::comma, limits = c(0, max_y * 1.2)) +
  labs(y = "Total number of Confirmed Covid-19 cases", x = "") +
  theme_classic()+
  coord_flip() 
```



```{r}
daily_cases_total %>% 
  filter(Type == "Confirmed") %>% 
  arrange(-Count_change) %>% 
  mutate(Country = forcats::fct_reorder(Country, Count_change))%>% 
  slice_head(n = top) %>% 
  ggplot(aes(x = Country, y = Count_lastest, 
             label = paste0(round(Count_lastest/1e6,1), "M", 
                            "(+", round(Count_change/1e3, 1), "K)"))) +
  geom_col() +
  geom_text(hjust = 0, fontface = "bold") + 
  scale_y_continuous(labels = scales::comma, limits = c(0, max_y * 1.2)) +
  labs(y = "Total number of Confirmed Covid-19 cases", x = "") +
  theme_classic()+
  coord_flip() 
```



## Active Cases

```{r, echo=FALSE}
max_y <-
  daily_cases_total %>%
  filter(Type == "Active") %>% pull(Count_lastest) %>% max(na.rm = TRUE)

daily_cases_total %>% 
  filter(Type == "Active") %>% 
  arrange(-Count_lastest) %>% 
  mutate(Country = forcats::fct_reorder(Country, Count_lastest))%>% 
  slice_head(n = top) %>% 
  ggplot(aes(x = Country, y = Count_lastest, 
             label = paste0(round(Count_lastest/1e6,1), "M", 
                            "(+", round(Count_change/1e3, 1), "K)"))) +
  geom_col() +
  geom_text(hjust = 0, fontface = "bold") + 
  scale_y_continuous(labels = scales::comma, limits = c(0, max_y * 1.2)) +
  labs(y = "Total number of Active Covid-19 cases", x = "") +
  theme_classic()+
  coord_flip() 
```


```{r, echo=FALSE}
daily_cases_total %>% 
  filter(Type == "Active") %>% 
  arrange(-Count_change) %>% 
  mutate(Country = forcats::fct_reorder(Country, Count_change))%>% 
  slice_head(n = top) %>% 
  ggplot(aes(x = Country, y = Count_lastest, 
             label = paste0(round(Count_lastest/1e6,1), "M", 
                            "(+", round(Count_change/1e3, 1), "K)"))) +
  geom_col() +
  geom_text(hjust = 0, fontface = "bold") + 
  scale_y_continuous(labels = scales::comma, limits = c(0, max_y * 1.2)) +
  labs(y = "Total number of Confirmed Covid-19 cases", x = "") +
  theme_classic()+
  coord_flip() 
```


## Deaths

```{r, echo=FALSE}
max_y <-
  daily_cases_total %>%
  filter(Type == "Deaths") %>% pull(Count_lastest) %>% max(na.rm = TRUE)

daily_cases_total %>% 
  filter(Type == "Deaths") %>% 
  arrange(-Count_lastest) %>% 
  mutate(Country = forcats::fct_reorder(Country, Count_lastest))%>% 
  slice_head(n = top) %>% 
  ggplot(aes(x = Country, y = Count_lastest, 
             label = paste0(round(Count_lastest/1e3,1), "K", 
                            "(+", Count_change, ")"))) +
  geom_col() +
  geom_text(hjust = 0, fontface = "bold") + 
  scale_y_continuous(labels = scales::comma, limits = c(0, max_y * 1.2)) +
  labs(y = "Total number of Covid-19 Deaths", x = "") +
  theme_classic()+
  coord_flip() 
```


```{r, echo=FALSE}
daily_cases_total %>% 
  filter(Type == "Deaths") %>% 
  arrange(-Count_change) %>% 
  mutate(Country = forcats::fct_reorder(Country, Count_change))%>% 
  slice_head(n = top) %>% 
  ggplot(aes(x = Country, y = Count_lastest, 
             label = paste0(round(Count_lastest/1e3,1), "K", 
                            "(+", Count_change, ")"))) +
  geom_col() +
  geom_text(hjust = 0, fontface = "bold") + 
  scale_y_continuous(labels = scales::comma, limits = c(0, max_y * 1.2)) +
  labs(y = "Total number of Covid-19 Deaths", x = "") +
  theme_classic()+
  coord_flip() 
```


## Recovered

```{r, echo=FALSE}

max_y <-
  daily_cases_total %>%
  filter(Type == "Recovered") %>% pull(Count_lastest) %>% max(na.rm = TRUE)

daily_cases_total %>% 
  filter(Type == "Recovered") %>% 
  arrange(-Count_lastest) %>% 
  mutate(Country = forcats::fct_reorder(Country, Count_lastest))%>% 
  slice_head(n = top) %>% 
  ggplot(aes(x = Country, y = Count_lastest, 
             label = paste0(round(Count_lastest/1e6,1), "M", 
                            "(+", round(Count_change/1e3, 1), "K)"))) +
  geom_col() +
  geom_text(hjust = 0, fontface = "bold") + 
  scale_y_continuous(labels = scales::comma, limits = c(0, max_y * 1.2)) +
  labs(y = "Total number of Covid-19 Recovered", x = "") +
  theme_classic()+
  coord_flip() 

```


```{r, echo=FALSE}
daily_cases_total %>% 
  filter(Type == "Recovered") %>% 
  arrange(-Count_change) %>% 
  mutate(Country = forcats::fct_reorder(Country, Count_change))%>% 
  slice_head(n = top) %>% 
  ggplot(aes(x = Country, y = Count_lastest, 
             label = paste0(round(Count_lastest/1e6,1), "M", 
                            "(+", round(Count_change/1e3, 1), "K)"))) +
  geom_col() +
  geom_text(hjust = 0, fontface = "bold") + 
  scale_y_continuous(labels = scales::comma, limits = c(0, max_y * 1.2)) +
  labs(y = "Total number of Covid-19 Recovered", x = "") +
  theme_classic()+
  coord_flip() 
```

